---
title: "Data Creation"
author: "Jes Stevens, M.A."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
always_allow_html: yes
params: 
  month: August 
  pastyear: 2018
  year: 2019 
  date: "2019-08-01"
  pastdate: "2018-01-01"
  pstdtmonth: "2018-08-01"
  pstdtmonth2: "2018-08-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

[Washington University Medical Center Redevelopment Corporation](http://wumcrc.com) is a partnership between BJC Health Care and Washington University School of Medicine and works to improve the quality of life for the neighborhoods surrounding the medical campus. In order to achieve this goal in Forest Park Southeast and the Central West End , WUMCRC has invested millions of dollars toward regenerating the market for private investment in businesses and real estate, enhancing human and social service opportunities, and improving the level of physical and personal security.

One way we work to improve the level of physical & personal security is the analysis and distribution of data. The original source of this crime data is <http://slmpd.org/crimereports.shtml>. This notebook uses primarily `compstatr` to access and clean the crime data. 

## R Markdown

```{r Load Dependencies, include = FALSE}
# tidyverse packages
library(ggplot2)       # plotting data
library(stringr)       # wrappers for common string operations
library(tidyr)         # tidy data
library(dplyr)         # data manipulation
library(magrittr)      # pipe operator
library(readxl)        # read & write excel files
library(lubridate)     # time data manipulation

# spatial packages
library(tmap)         # map layouts
library(tmaptools)    # tools for handeling spatial 
library(oldtmaptools) # deprecated toolds for spatial analysis
library(sf)           # spatial data tools
library(ceramic)      # download online imagery tiles
library(compstatr)    # tools for STL crime data
library(raster)       # geograpic data analysis & modeling

# other packages
library(here)         # file path management
library(janitor)      # tools for examining data

dst2 <- c(7:15,27:29, 39:45,81,82,87,88)
dst5 <- c(38,46:58,78)

load(here("data/nbhd_pop10.rda"))
```

## Tidy data
### Get Crime Data

```{r Pull 2018 and 2019 Data}
i <- cs_create_index()
yearlist18 <- cs_get_data(year = params$pastyear, index = i)
yearlist19 <- cs_get_data(year = params$year, index = i) # Constructs a table for finding a table of crime data
cs_validate(yearlist18, year = 2018)
cs_validate(yearlist19, year = 2019) # Identifies problems with SLMPD Data
```

#### Standardize Data for Cleaning

```{r Standardize 2018 and 2019 Data}
totalCrimes18 <- cs_collapse(yearlist18)
ytdCrimes19 <- cs_collapse(yearlist19) 
```

### Clean & Categorize Data 
#### 2019

`cs_filter_count` removes negative counts. Negative counts, -1, in the count column means that the crime, or charge in this specific observation has either been deemed unfounded, or the crime has been up coded. We do not want to map this data. 

Many of the analyses we conduct include comparisons between violent & non-violent crime, comparisons on the amount of crimes happening in each crime cateogy over time, and if crimes occur during the day or at night. The following code ceates variables to conduct these analyses. 

`cs_crime_cat` creates a variable with the names of the crime. 

`cs_crime` creates a logic variable and codes categories of crimes as either `TRUE` or `FALSE` based on the input.

`cs_parse_date` creates two columns separating the `Date Occur` variable. The two colums are as follows: one contains the date - month, date, and year, and the other contains the hour and minute. This is used because crimes coded in the most recent month, can contain dates that occured, in previous months or years & in this report we only want to map the crimes that occured in the past month. 

`filter` is a `dplyr` function that filters out any dates that occur before the our selected date, and also filters out crimes that did not happen in either District 2 or district 5. 

`mutate` adds a variable that codes and labels the days of the week for each crime that occurred, and creates another time of day variable. 

`strptime` and `format` takes the new time variable and formats it to a character so that we can determine if the crime occured at day or at night, and creates a second coded variable that labels each observations as day or night based on the newly formated time variable. 

`select` drops the unneeded variables. 

`cs_missing_XY` determines what data does not have x & y coordinates, and therefore cannot be accurately mapped. 

`cs_replace0` replaces missing x & y coordinates with `NA`, and drops the missing data. 

`strptime` and `format` takes the new time variable and formats it to a character so that we can determine if the crime occured at day or at night, and creates a second coded variable that labels each observations as day or night based on the newly formated time variable. 

`select` drops the unneeded variables. 

`cs_missing_XY` determines what data does not have x & y coordinates, and therefore cannot be accurately mapped. 

`cs_replace0` replaces missing x & y coordinates with `NA`, and drops the missing data. 

`monthCrimes` is an object that holds crimes for the current month of this year. 

```{r Creates 2019 Tidy Crime File}
ytdCrimes19 <- ytdCrimes19 %>% 
  cs_filter_count(., var = count) %>%
  cs_filter_crime(., var = crime, "part 1") %>%
  cs_crime_cat(., var = crime, crimeCatNum, "numeric") %>%
  cs_crime_cat(., var = crime, crimeCatName, "string") %>%
  cs_crime(., var = crime, violent, "violent") %>%
  cs_crime(., var = crime, property, "property") %>%
  cs_parse_date(., date_occur, dateVar = dateOcc, timeVar = timeOcc) %>%
  filter(district == 2 | district == 5) %>%
  mutate(weekday = wday(dateOcc, label = TRUE)) %>%
  mutate(monthVar = dateOcc) %>%
  mutate(tod = timeOcc)

ytdCrimes19$monthVar <- month(as.Date(ytdCrimes19$monthVar, format="%d/%m/%Y"))

ytdCrimes19$neighborhood <- as.numeric(ytdCrimes19$neighborhood)
ytdCrimes19$tod <- strptime(ytdCrimes19$tod, tz = "America/Chicago", "%H:%M")
ytdCrimes19$tod <- format(ytdCrimes19$tod, format = "%H%M%S")

ytdCrimes19 <- ytdCrimes19 %>%
  mutate(., dayNight = ifelse(tod >= "180000" & tod < "600000", "Night", "Day")) %>% 
  dplyr::select(-dateTime, -tod, -flag_crime, -flag_administrative, -flag_unfounded, -flag_cleanup)  %>%
  cs_missingXY(., varX = x_coord, varY = y_coord, newVar = missing) %>% 
  cs_replace0(., var = x_coord) %>%
  cs_replace0(., var = y_coord) %>% 
  filter(., missing == FALSE) 

monthCrimes19 <- ytdCrimes19 %>% 
    filter(dateOcc >= as.Date(params$date))

rm(yearlist19)
```

##### Larcenies 

```{r}
larceny <- monthCrimes19 %>% 
  filter(., crimeCatNum == 6)

larceny <- mutate(larceny, "type" = ifelse(grepl("ALL OTHER", description), "All Other", 
                                    ifelse(grepl("BICYCLE", description), "Bicycle", 
                                    ifelse(grepl("FROM BUILDING", description), "From Building", 
                                    ifelse(grepl("FROM COIN", description), "From Coin Machine", 
                                    ifelse(grepl("FROM MTR VEH", description), "From Motor Vehicle", 
                                    ifelse(grepl("MTR VEH PARTS", description), "Motor Vehicle Parts",
                                    ifelse(grepl("PURSE", description), "Purse Snatching",
                                    ifelse(grepl("SHOPLIFT", description), "Shoplifting",
                                    ifelse(grepl("FRM PRSN", description), "From Person", NA))))))))))

larceny <- mutate(larceny, "value" = ifelse(grepl("UNDER", description), "Under $500",
                                     ifelse(grepl("\\$500 - \\$24,999", description), "$500 - $24,999",
                                     ifelse(grepl("OVER \\$25,000", description), "Over $25,000", NA)))) 
```

#### 2018

```{r Total Crimes for 2018 Tidied}
totalCrimes18 <- totalCrimes18 %>% 
  cs_filter_count(., var = count) %>%
  cs_filter_crime(., var = crime, "part 1") %>%
  cs_crime_cat(., var = crime, crimeCatNum, "numeric") %>%
  cs_crime_cat(., var = crime, crimeCatName, "string") %>%
  cs_crime(., var = crime, violent, "violent") %>%
  cs_crime(., var = crime, property, "property") %>%
  cs_parse_date(., date_occur, dateVar = dateOcc, timeVar = timeOcc) %>%
  filter(district == 2 | district == 5) %>%
  mutate(weekday = wday(dateOcc, label = TRUE)) %>%
  mutate(monthVar = dateOcc) %>%
  mutate(tod = timeOcc) 

totalCrimes18$monthVar <- month(as.Date(totalCrimes18$monthVar, format="%d/%m/%Y"))

totalCrimes18$neighborhood <- as.numeric(totalCrimes18$neighborhood)

totalCrimes18$tod <- strptime(totalCrimes18$tod, tz = "America/Chicago", "%H:%M")
totalCrimes18$tod <- format(totalCrimes18$tod, format = "%H%M%S")

totalCrimes18 <- totalCrimes18 %>%
  mutate(., dayNight = ifelse(tod >= "180000" & tod < "600000", "Night", "Day")) %>% 
  dplyr::select(-dateTime, -tod, -flag_crime, -flag_administrative, -flag_unfounded, -flag_cleanup) %>%
  cs_missingXY(., varX = x_coord, varY = y_coord, newVar = missing) %>%
  cs_replace0(., var = x_coord) %>%
  cs_replace0(., var = y_coord) %>% 
  filter(., missing == FALSE) %>% 
  filter(dateOcc >= as.Date(params$pastdate))

monthCrimes18 <- totalCrimes18 %>% 
    filter(dateOcc >= as.Date(params$pstdtmonth) & dateOcc <= as.Date(params$pstdtmonth2))

ytdCrimes18 <- totalCrimes18%>% 
    filter(dateOcc <= as.Date(params$pstdtmonth2))

rm(yearlist18)
```


### Get Spatial Data 

#### External Shapefiles 

```{r Load Shapefiles, include=FALSE}
st_read(here("data", "external", "med-campus", "medcampus.shp")) %>%
  st_transform(crs = 102696) -> med_campus

st_read(here("data", "external", "cortex-boundary", "CORTEX BOUNDARY.shp")) %>%
  st_transform(crs = 102696) -> cortex

st_read(here("data", "external", "grove-cid", "grovecid.shp")) %>%
  st_transform(crs = 102696) -> grove_cid

st_read(here("data", "external", "stl_neighborhoods", "nbhds.shp")) %>%
  st_transform(crs = 102696) %>% 
  rename(., "neighborhood" = NHD_NUM)-> nhoods_sf

save(med_campus, cortex, grove_cid, nhoods_sf, file = here("data", "basemap-files", "boundaries.rda"))
```

#### Tiles from Mapbox 

```{r Service Area Neighborhood Coordinates and Basemap Tiles}
xyfpse <- c(-90.2679, -90.2423, 38.6176, 38.6334)
xycwe <- c(-90.2759, -90.2368, 38.6286, 38.6552)
xybot <- c(-90.2619, -90.2409, 38.6165, 38.6296)
xydbp <- c(-90.2869, -90.2726, 38.6433, 38.6566)
xysdb <- c(-90.3026, -90.2827, 38.6456, 38.6571)
xywe <- c(-90.3020, -90.2712, 38.6517, 38.6710)
xyvp <- c(-90.2803, -90.2712, 38.6517, 38.6622)
xyac <- c(-90.2744, -90.2609, 38.6505, 38.6661)
xyfp <- c(-90.2648, -90.2543, 38.6493, 38.6655)
xylp <- c(-90.2588, -90.2437, 38.6481, 38.6624)
xyvd <- c(-90.2520, -90.2304, 38.6426, 38.6585)
xymc <- c(-90.2678, -90.2515, 38.6305, 38.6411)
xyctx <- c(-90.2581, -90.2419, 38.6299, 38.6386)
xygrv <- c(-90.2662, -90.2440, 38.6238, 38.6318)
xydst2 <- c(-90.3203, -90.2297, 38.5613, 38.6493)
xydst5 <- c(-90.3080, -90.2132, 38.6273, 38.6962)

fpse_tiles <- raster::extent(xyfpse) %>%
  cc_location(., type = "mapbox.streets", max_tiles = 15)

cwe_tiles <- raster::extent(xycwe) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

bot_tiles <- raster::extent(xybot) %>%
  cc_location(., type = "mapbox.streets", max_tiles = 15)

dbp_tiles <- raster::extent(xydbp) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

sdb_tiles <- raster::extent(xysdb) %>%
  cc_location(., type = "mapbox.streets", max_tiles = 15)

we_tiles <- raster::extent(xywe) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

vp_tiles <- raster::extent(xyvp) %>%
  cc_location(., type = "mapbox.streets", max_tiles = 15)

ac_tiles <- raster::extent(xyac) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

fp_tiles <- raster::extent(xyfp) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

lp_tiles <- raster::extent(xylp) %>%
  cc_location(., type = "mapbox.streets", max_tiles = 15)

vd_tiles <- raster::extent(xyvd) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

mc_tiles <- raster::extent(xymc) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

ctx_tiles <- raster::extent(xyctx) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

grv_tiles <- raster::extent(xygrv) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

dst2_tiles <- raster::extent(xydst2) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

dst5_tiles <- raster::extent(xydst5) %>% 
  cc_location(., type = "mapbox.streets", max_tiles = 15)

save(fpse_tiles, cwe_tiles, bot_tiles, dbp_tiles, sdb_tiles, we_tiles, vp_tiles, ac_tiles, fp_tiles, lp_tiles, vd_tiles, mc_tiles, ctx_tiles, grv_tiles, dst2_tiles, dst5_tiles, file = here("data", "basemap-files", "mapbox-tiles.rda"))

fpse <- filter(nhoods_sf, neighborhood == 39 )
cwe <- filter(nhoods_sf, neighborhood == 38 )
bot <- filter(nhoods_sf, neighborhood == 28 )
dbp <- filter(nhoods_sf, neighborhood == 47 )
sdb <- filter(nhoods_sf, neighborhood == 46 )
we <- filter(nhoods_sf, neighborhood == 48 )
vp <- filter(nhoods_sf, neighborhood == 49 )
ac <- filter(nhoods_sf, neighborhood == 51 )
fp <- filter(nhoods_sf, neighborhood == 53 )
lp <- filter(nhoods_sf, neighborhood == 54 )
vd <- filter(nhoods_sf, neighborhood == 58 )

save(fpse, cwe, bot, dbp, sdb, we, vp, ac, fp, lp, vd, file = here("data", "basemap-files", "nbhd-boundaries.rda"))
```

```{r Create SF objects for 2018 & 2019 Crimes}
# all crimes 
totalCrimes18_sf <- cs_projectXY(totalCrimes18, varX = x_coord, varY = y_coord, crs = 102696)
ytdCrimes18_sf <- cs_projectXY(ytdCrimes18, varX = x_coord, varY = y_coord, crs = 102696)
ytdCrimes19_sf <- cs_projectXY(ytdCrimes19, varX = x_coord, varY = y_coord, crs = 102696)

monthCrimes18_sf <- cs_projectXY(monthCrimes18, varX = x_coord, varY = y_coord, crs = 102696)
monthCrimes19_sf <- cs_projectXY(monthCrimes19, varX = x_coord, varY = y_coord, crs = 102696)

larceny_sf <- cs_projectXY(larceny, varX = x_coord, varY = y_coord, crs = 102696)

# med campus
mc_crimes_month18_sf <- st_intersection(monthCrimes18_sf, med_campus) 
mc_crimes_month18 <- mc_crimes_month18_sf  %>% 
   as.data.frame() 

mc_crimes_month19_sf <- st_intersection(monthCrimes19_sf, med_campus)
mc_crimes_month19 <- mc_crimes_month19_sf  %>% 
   as.data.frame()

mc_crimes_ytd18_sf <- st_intersection(ytdCrimes18_sf, med_campus) 
mc_crimes_ytd18 <- mc_crimes_ytd18_sf %>% 
   as.data.frame() 

mc_crimes_ytd19_sf <- st_intersection(ytdCrimes19_sf, med_campus)
mc_crimes_ytd19 <- mc_crimes_ytd19_sf %>% 
   as.data.frame()

mc_crimes_total18_sf <- st_intersection(totalCrimes18_sf, med_campus)
mc_crimes_total18 <- mc_crimes_total18_sf %>% 
   as.data.frame()

mc_larcenies_sf <- st_intersection(larceny_sf, med_campus)
mc_larcenies <- mc_larcenies_sf %>% 
   as.data.frame()

# cortex
ctx_crimes_sf <- st_intersection(monthCrimes19_sf, cortex)

# police districts 
dst_2 <- monthCrimes19 %>% 
  filter(., neighborhood %in% dst2) %>% 
  group_by(., neighborhood) %>%
  count() %>% 
  rename(crimeTotal = n) %>%
  left_join(nbhd_pop10, by = "neighborhood") %>% 
  mutate(., crimeRate = (crimeTotal/pop10)*1000) %>% 
  drop_na()

dst_5 <- monthCrimes19 %>% 
  filter(., neighborhood %in% dst5) %>% 
  group_by(., neighborhood) %>%
  count() %>% 
  rename(crimeTotal = n) %>%
  left_join(nbhd_pop10, by = "neighborhood") %>% 
  mutate(., crimeRate = (crimeTotal/pop10)*1000) %>% 
  drop_na()

dst_2_pop_sf <- left_join(nhoods_sf, dst_2, by = "neighborhood") %>% 
  st_transform(crs = 102696) %>%
  drop_na() %>% 
  subset(., neighborhood != 88)

dst_5_pop_sf <- left_join(nhoods_sf, dst_5, by = "neighborhood") %>% 
  st_transform(crs = 102696) %>%
  drop_na()

# Save
save(totalCrimes18_sf, ytdCrimes18_sf, ytdCrimes19_sf, monthCrimes18_sf, monthCrimes19_sf, mc_crimes_month18_sf, mc_crimes_month19_sf, mc_crimes_ytd18_sf, mc_crimes_total18_sf, mc_crimes_ytd19_sf, mc_larcenies_sf, ctx_crimes_sf, dst_2_pop_sf, dst_5_pop_sf, file = here("data", "spatial-crime-data.rda"))

save(ytdCrimes18, ytdCrimes19, monthCrimes18, monthCrimes19, totalCrimes18, larceny, mc_crimes_month18, mc_crimes_month19, mc_crimes_ytd18, mc_crimes_total18, mc_crimes_ytd19, mc_larcenies, dst_2, dst_5, file = here("data", "crime-data.rda"))

rm(list = ls())
```
